Kepler's Demographic Cliff: Ages of Young Stars and Planets in the Kepler Field

### 2024-05-21 paper/ms.tex
update ms

@@

Stars that formed in the same birth cluster 
are the gold standard for the astronomical age scale
\citep{Soderblom_2010}.
Before Gaia, the few $\gtrsim$1\,Gyr open clusters
known in the Kepler field had been visually identified by
\citet{1864RSPT..154....1H}.
Gaia proper motions and distances 
enabled the discovery of stellar ensembles that are typically more diffuse,
but which nonetheless
share a common age 
based on e.g.~isochrone, rotation, and lithium dating
\cite[e.g.][]{2019AJ....158..122K,Bouma_2022b,Barber_2022}.

In Figure~\ref{fig:agescalecompone}, we compare 
ages from the literature against the ages of stars in the richest
clusters in the Kepler field.
In particular, we drew ages for Kepler stars from \citet{Berger_2020a_catalog},
\citet{2021AJ....161..189L}, \citet{2023ApJ...952..131M}, and
\citet{2024AJ....167..159L}.
We followed any guidance available from these respective studies for
removing ages that were explicitly flagged as unreliable.

@@

The point of Figure~\ref{fig:agescalecompone} is hopefully clear.
While previous studies have been able to derive accurate ages for
stars $\gtrsim$1\,Gyr old, the ages of sub-gigayear stars have historically been
overestimated by anywhere from 0.5--2\,dex.
The situation is particularly severe for the 40\,Myr stars in Cep-Her,
but it extends out to 150--300\,Myr stars in Melange-3 and Theia~520.
For isochrone and kinematic ages,
this is not surprising because these methods use stellar parameters that do
not appreciably change at $t$$\lesssim$1\,Gyr.
For the rotation-based ages quoted by
\citet{2023ApJ...952..131M} and \citet{2024AJ....167..159L}, the discrepancy
between their ages and the cluster age scale is likely caused by details of the calibration 
methodologies.
The physical spin-down model used by
\citet{2023ApJ...952..131M},
and the gyro-kinematic calibration ages used by \citet{2024AJ....167..159L},
may not reproduce
the rotation sequences of open clusters such as Praesepe, the Hyades, and the Pleiades.
The statistical age uncertainties yielded by these approaches
similarly may not marginalize over the intrinsic
scatter in rotation periods present in stars of a fixed mass at any given age.

>>>>

% Redefine footnote command to not have indent
%\makeatletter
%\long\def@makefntext#1{\noindent\footnotesize@makefnmark#1}
%\makeatother
\setlength{\footnotesep}{0pt}

@@

Stars that formed in the same birth cluster 
are the gold standard for the astronomical age scale
\citep{Soderblom_2010}.
Before Gaia, the few $\gtrsim$1\,Gyr open clusters
known in the Kepler field had been visually identified by
\citet{1864RSPT..154....1H}.
Gaia proper motions and distances 
enabled the discovery of stellar ensembles that are more diffuse,
but which nonetheless
share a common age 
based on e.g.~isochrone, rotation, and lithium dating
\cite[e.g.][]{2019AJ....158..122K,Bouma_2022b,Barber_2022}.

In Figure~\ref{fig:agescalecompone}, we compare 
ages from the literature against the ages of stars in the richest
clusters in the Kepler field.
In particular, we drew ages for Kepler stars from \citet{Berger_2020a_catalog},
\citet{2021AJ....161..189L}, \citet{2023ApJ...952..131M}, and
\citet{2024AJ....167..159L}.
We followed any guidance available from these respective studies for
removing ages that were explicitly flagged as unreliable, and plotted
stars within $1\sigma$ of zero as being upper limits.

@@

The point of Figure~\ref{fig:agescalecompone} is hopefully clear.
While previous studies have been able to derive accurate ages for
stars $\gtrsim$1\,Gyr old, the ages of sub-gigayear stars have historically been
overestimated by anywhere from 0.5--2\,dex.
The situation is particularly severe for the 40\,Myr stars in Cep-Her,
but it extends out to 150--300\,Myr stars in Melange-3 and Theia~520.
For isochrone and kinematic ages,
this is not surprising because these methods use stellar parameters that do
not appreciably change at $t$$\lesssim$1\,Gyr.
For the rotation-based ages quoted by
\citet{2023ApJ...952..131M} and \citet{2024AJ....167..159L}, the discrepancy
between their ages and the cluster age scale is likely caused by details of the calibration 
methodologies.
The physical spin-down models used by
\citet{2023ApJ...952..131M},
and the gyro-kinematic calibration ages used by \citet{2024AJ....167..159L},
may not reproduce
the rotation sequences of open clusters such as Praesepe, the Hyades, and the Pleiades.
The statistical age uncertainties yielded by these approaches
similarly may not marginalize over the intrinsic
scatter in rotation periods present in stars of a fixed mass at any given age.
